---
title: "Data Storage and Retrieval -- Assignment 11"
author: "Aaron Niskin"
date: "November 18, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r, echo=FALSE}
setwd("~/Documents/courses/newCollege/current/databases/assignments/a11/")
```



1.  **Do Problem 1. See the group() finalize example on page 157.**
    
    ```javascript
    // Returns an object of prefix: count_for_prefix
    db.phones.group({
      initial: {prefixes : {}},
      reduce: function(phone, output) {
        var out = output.prefixes,
            pre = phone.components.prefix;
        if (out.hasOwnProperty(pre)){
          out[pre] += 1;
        } else {
          out[pre] = 1;
        }
      },
      finalize: function(out) {
        return out;
      }
    })[0].prefixes
    
    // Returns the count of unique prefixes
    db.phones.group({
      initial: { prefixes : {} },
      reduce: function(phone, output) {
        output.prefixes[phone.components.prefix] = 1;
      },
      finalize: function(out) {
        out.total = Object.keys(out.prefixes).length
      }
    })[0].total
    ```
    
2.  **Do Problem 2. Create your own database. Use python and the countries.json collection in /usr/share/databases/SevenDatabases/code/mongo/. Create a text index on the common name. Perform a query using the $text operator and show the results.**
    
    **Install a Mongo driver for a language of your choice, and connect to the database. Populate a collection through it, and index one of the fields.**
    
    ```python
    ### See the code file: /usr/share/databases/aniskin/assignment11/q2.py
    ### and the accompanying results json file: /usr/share/databases/aniskin/assignment11/yentel.json
    import pymongo
    from bson.json_util import dumps
    import json
    from pprint import PrettyPrinter
    pp = PrettyPrinter(indent = 2)
    
    client = pymongo.MongoClient("localhost", 27017)
    db = client.countries_aaron
    
    db.countries_of_adawg.drop()
    with open("/usr/share/databases/SevenDatabases/code/mongo/countries.json") as json_file:
            json_data = json.load(json_file)
    
    ids = [db.countries_of_adawg.insert(x) for x in json_data]
    
    db.countries_of_adawg.create_index([("name.common", 'text')], default_language='english')
    
    kibbitzer = db.countries_of_adawg.find_one({"$text": { "$search": "Israel" } })
    pp.pprint(kibbitzer)
    
    with open('yentel.json', 'w') as kvetcher1:
            json.dump(json.loads(dumps(kibbitzer)), kvetcher1)
    
    print("You're leaving? But I just made a meal. No, no... It's fine. I guess I can eat it all myself. I'll get fat. It's okay. You know, I slave over the hot stove for hours, but no, it's okay.")
    ```
    
3.  **For the country collection you created in problem 2,**
    
    a.  **Use group() to compute the area of each region**
        
        ```javascript
        db.countries_of_adawg.group({
          initial: {},
          reduce: function(country, out) {
            var region = country["region"],
                area = country["area"];
            if(out.hasOwnProperty(region)) {
              out[region] += parseInt(area);
            } else {
              out[region] = parseInt(area);
            }
          },
          finalize: function(out) {
            return out;
          }
        })[0]
        ```
        
        {
          
            "Americas" : 42081497,
            
            "Asia" : 32138141,
            
            "Africa" : 30318023,
            
            "Europe" : 23022897,
            
            "Oceania" : 8515313,
            
            "" : 14008208
          
        }

    
    b.  **Use MapReduce to find the number of countries that border N other countries for each N.**
    
        ```javascript
         db.countries_of_adawg.group({
          initial: {},
          reduce: function(country, out) {
            var borders = country.borders.length;
            if(out.hasOwnProperty(borders)) {
              out[borders] += 1;
            } else {
              out[borders] = 1;
            }
          },
          finalize: function(out) {
            return out;
          }
        })[0]
        ```
        {
        	"0" : 82,
        	"1" : 24,
        	"2" : 28,
        	"3" : 30,
        	"4" : 26,
        	"5" : 24,
        	"6" : 13,
        	"7" : 9,
        	"8" : 7,
        	"9" : 2,
        	"10" : 1,
        	"14" : 1,
        	"15" : 1
        }
