---
title: "Data Storage and Retrieval -- Assignment 10"
author: "Aaron Niskin"
date: "November 07, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
setwd("~/Documents/courses/newCollege/current/databases/assignments/a10/")
```



1.  **Print a JSON document containing `{"hello": "world"}`**
    
    ```javascript
    printjson({"hello": "world"})
    ```
    
2.  **Select a town via a case-insensitive regular expression containing the word *new*.**
    
    ```javascript
    db.cities.find({name: /new/i}).limit(1)
    ```
    
3.  **Find all cities whose names contain *e* and are famous for food or beer.**
    
    ```javascript
    db.towns.find({name: /e/i, famous_for: {$in: ["food", "beer"]}})
    ```

4.  **Create a new database named *blogger* with a collection named *articles*. Insert a new article with an author name and email, creation date, and text.**
    
    ```javascript
    use blogger;
    db.articles.insert(
    {author:
      {name: "Steve I-cant-spell Buschemi",
       email: "StevieWonderful@hotlink.com"},
       creation_date: "yesterday (that won't ever be confusing)",
       text: "Has he ever been in a bad movie? How has he not gotten an Oscar yet? Wait... Has he gotten an Oscar? I guess this is why I need internet at my apartment. But nope. Won't deal with Comcast. And I think that's a long enough comment. I need sleep."})
    ```

5.  **Update the article with an array of comments, containing a comment with an author and text.**
    
    ```javascript
    db.articles.update(
      {"author.name": /Steve/i},
      {$set: {comments: 
              [{author: "John Clease", text: "We are the nights who say, NI!"},
               {author: "Jack Doneghy", text: "It's spelled with a 'k', buddy."}
              ]}})
    ```
    
6.  **Run a query from external JavaScript and Python files.**
    
    This isn't really something I can put here, but I'll try.
    
    ```bash
    mongo cities < tmp.js
    ```
    
    Where `tmp.js` is normal JavaScript code one might execute from within Mongo. In my case, it's just `db.cities.find().limit(1)` (both the collection and the database are called `cities`, unfortunately). As for Python, we will simply execute the code directly.
    
    ```{python}
    from pymongo import MongoClient
    mg = MongoClient()
    print(mg.database_names())
    ```
